---
title:           How to make graph percentage scaled?
author:          Dennis Wollersheim & Josh Vartuli
date:            11.06.2020
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---


There are many ways to produce scales.  Here are a few.

# load libraries

```{r}

library(tidyverse)
library(RPostgreSQL)

```

# Get data

```{r}

con <- dbConnect(dbDriver("PostgreSQL"),
                 dbname   = 'vaed_full',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query_diag <- "
select admission_id, diag_code, diag_desc, sex_desc
from admission
JOIN sex using (sex_id)
join admission_diagnosis using (admission_id)
join diagnosis_desc using (diag_code)
  WHERE diag_desc like '%leukaemia%' ;
  "

leukaemia_diag <- dbGetQuery( con, query_diag ) %>% as_tibble()
did_disconnect <- dbDisconnect( con )
  leukaemia_diag %>%
distinct(admission_id)

```
# Original Query
The problem with this query is that it is plotting admission_id as a numeric variable, leading to values of 2 billion!

```{r}

  leukaemia_diag%>%
  ggplot()+
  geom_col(mapping=aes(x=diag_desc, y=admission_id, color=sex_desc)) +
  ggtitle('leukaemia by gender and type') +
  xlab ('Diagnosis') +
  ylab ('Admissions') +
  scale_fill_discrete ('sex_desc') +
  theme(legend.position = 'bottom')


```

# Original Query - fixed

I imagine that you want to count diag_desc (and use fill instead of color).  The most simple way to do this is to use geom_bar do to the counting for us

```{r}

leukaemia_diag%>%
  ggplot()+
  geom_bar(mapping=aes(x=diag_desc, fill=sex_desc)) +
  ggtitle('leukaemia by gender and type') +
  xlab ('Diagnosis') +
  ylab ('Admissions') +
  scale_fill_discrete ('sex_desc') +
  theme(legend.position = 'bottom')

```

# How to produce percentage scales

## 1) Calculate the percentage yourself

*count* counts the number of admissions with each *diag_desc*, and *mutate* calculates the percentage.  We also can use *geom_co*l because we already know how high the bars should be, we don't need to have *geom_bar* do the counting for us.

```{r}

leukaemia_diag %>%
  count( diag_desc, sex_desc ) %>%
  mutate( percentage = n / sum(n) ) %>%
  ggplot()+
  geom_col(mapping=aes(x=diag_desc, y=percentage, fill=sex_desc)) +
  ggtitle('leukaemia by gender and type') +
  xlab ('Diagnosis') +
  ylab ('Admissions') +
  scale_fill_discrete ('sex_desc') +
  theme(legend.position = 'bottom')

```


## 2) Use magic word *..count..* in *geom_bar*

As you may imagine, because *geom_bar* can do the counting for us, it actually does know the count!  We can see and use the count, because it is stored in the *..count..* variable.


```{r}

leukaemia_diag %>%
  ggplot()+
  geom_bar(mapping=aes(x=diag_desc, y= ..count.. / sum( ..count..), fill=sex_desc)) +
  ggtitle('leukaemia by gender and type') +
  xlab ('Diagnosis') +
  ylab ('Admissions') +
  scale_fill_discrete ('sex_desc') +
  theme(legend.position = 'bottom')

```

## 3) Labelling scales

We probably want to label the scales appropriately.  The *scales* package does this for us.

```{r}

library(scales)

leukaemia_diag %>%
  ggplot()+
  geom_bar(mapping=aes(x=diag_desc, y= ..count.. / sum( ..count..), fill=sex_desc)) +
  ggtitle('leukaemia by gender and type') +
  xlab ('Diagnosis') +
  ylab ('Admissions') +
  scale_fill_discrete ('sex_desc') +
  theme(legend.position = 'bottom') +
  scale_y_continuous(labels=scales::percent)

```

# Further work

There are some videos, and tutorial 6, on how to reduce the number of categories
