---
title:           How to approach obscure error messages
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          htmll
---
```{r message=F, show=F}

library(tidyverse)
library(readxl)
options(workflowr.sysgit='')


```


# The Question - 'non-numeric argument to binary operator'

I am trying to convert birth_year to age. ...  This is what I typed into R:
```{r}

firstFitted_ind_state <- read_excel("firstFitted_ind_state.xlsx")

# changing the column name from birth_year to age
colnames(firstFitted_ind_state)[colnames(firstFitted_ind_state)=="birth_year"] <-"age"

#
```

```{r eval=F}

firstFitted_ind_state <- firstFitted_ind_state %>%
  mutate(age = as.numeric(2017-age))

```

And the following error occurred:

`Error in 2017 - age : non-numeric argument to binary operator`

# Solution


## Renaming a field

Very good!

I like how you figured out how to rename the column name in your tibble from `birth_year` to `age`.  

` colnames(firstFitted_ind_state)[colnames(firstFitted_ind_state)=="birth_year"] <-"age"`

I like to use tidyverse and pipes for that, for example:

```{r}

read_excel("firstFitted_ind_state.xlsx")  %>%
  rename( age = birth_year )

```

## changing a field

Instead of changing `birth_date`, I would create an additional `age` field.  


```{r mutate, eval = F}

read_excel("firstFitted_ind_state.xlsx")  %>%
  mutate( age = 2017 - birth_year )


```
`Error: Evaluation error: non-numeric argument to binary operator.`

We get the same error.  The most important word here is `non-numeric`. We are trying to do something to the `birth_year` column.  Looking at the input tibble:

```{r display}

read_excel("firstFitted_ind_state.xlsx" )

```

`birth_year` is as a `<chr>` type  This is reason for the error, because you cannot do math with character values.   Convert `birth_date` to a numeric before calculating.

```{r}

read_excel("firstFitted_ind_state.xlsx")  %>%
  mutate( age = 2017 - as.numeric( birth_year ))

```

This creates your age column, but the warning points to the actual problem, the reason that the column was classified as character in the first place. There is some data that is not numeric.

## Searching for problematic values in `birth_date`

The first 10 `birth_year` all look like numbers, so the character data must be somewhere else.  We could `View` it to find the error, or look at the `distinct` values, or `arrange` it ascending or descending.

```{r error_check}

read_excel("firstFitted_ind_state.xlsx" ) %>%
  distinct( birth_year )

read_excel("firstFitted_ind_state.xlsx" ) %>%
  arrange( birth_year )


read_excel("firstFitted_ind_state.xlsx" ) %>%
  arrange( desc( birth_year ))

```
## Filtering out the problem rows

The last row, with the `Total` `birth_year` value, is causing the problem.  Let's filter it out, and try again:

```{r filter}

read_excel("firstFitted_ind_state.xlsx")  %>%
  filter( birth_year != 'Total' ) %>%
  mutate( age = 2017 - as.numeric( birth_year ))

```

Problem solved

