---
title:           How to approach obscure error messages
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          html
---


# Anthea's question about factors

Initially, I tried to identify admissions with HAC and without HAC by:

## load libraries
```{r libraries, echo=TRUE, warnings=FALSE, messages=FALSE}
# load libraries
library(tidyverse)
library("RPostgreSQL")

```
## Get data from SQL

```{r database_setup, echo=TRUE, warnings=FALSE, messages=FALSE}

# create connection to the database
con <- dbConnect(dbDriver("PostgreSQL"), 
                 dbname   = 'vaed_full',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query_diag <-"
SELECT  
    admission_id,
    sex_desc,
    age_years,
    los,
    sameday_id,
    care_type_id,
    care_type_desc,
    drg_description,
    diag_code,
    procedure_code,
    block
FROM admission
JOIN sex USING (sex_id)
JOIN admission_diagnosis USING (admission_id)
JOIN admission_procedure USING (admission_id)
JOIN procedure_desc USING (procedure_code)
JOIN diagnosis_desc USING (diag_code)
JOIN drg USING (drg)
JOIN care_type USING (care_type_id)
WHERE 
    block BETWEEN '672' and '679'
"

bypass <- dbGetQuery(con, query_diag) %>%
  as_tibble() %>%
  mutate(hac_status = ifelse(diag_code == 'Y832', 'HAC', 'No HAC')) 

```
  I count to confirm that it's correct:

```{r}

bypass %>%
  count(hac_status)

```

The results match PgAdmin using SELECT DISTINCT

But then... I was thinking, if R reads line by line, there will be lines where the admission_id remains the same but the code is not 'Y832' (Surgical operation with anastomosis, bypass or graft) as diag_code to admission_id is M:1. So I decided to assign `no HAC` to all admissions that fits in the admission_id that contains `Y832`.

(Note, I changed your query to use `bypass1` instead of `bypass` so I didn't overwrite the original bypass

```{r}

query_hac <-"
  SELECT admission_id
  FROM admission_diagnosis
  WHERE diag_code = 'Y832'
"
#
bypass1 <- dbGetQuery(con, query_hac) %>%
  as_tibble() %>%
  mutate(hac = ifelse(admission_id == query_hac, 'HAC', 'No HAC')) 

```

When I check with 

```{r}

bypass1 %>%
  count(hac)

```
It assigns all admissions as 'No HAC'
What happened here?

Anthea

# My answer

Let's investigate:  what is in bypass1?


```{r}

bypass1 


```
`query_hac` is a very simple query, it only retrieves `admission_id`, and you add in `hac` using `mutate` in the next line. The `mutate` condition depends on `admission_id` and `query_hac`, so, what is in `query_hac`?

```{r}
query_hac
```
`admission_id` is never equivalent to `query_hac`, so, it always says `No Hac`

I don't think that was your intention, but I am not sure what you are trying to do...

