---
title:           
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          html
---
```{r message=F, show=F}

library(tidyverse)
cardiac_diagnosis = read_csv("cardiac_diagnosis.csv") 


```


# Anthea's Original Question

>Hi Dennis
>
>I am trying to attempt the following question:
>
>For the most frequent care type, what is the gender breakdown of patients?
>
>
>I tried to select everything then filter by the most common care type, so I conjure the following:

```{r  eval=F}
 cardiac_diagnosis %>%
     distinct(admission_id, care_type_desc, sex_desc) %>%
     filter(
         count(care_type_desc, sort=TRUE) %>%
             head(1)
     )

```

>It made no sense in R.... of course. The following is the error message:

>Error in UseMethod("groups") : 
>no applicable method for 'groups' applied to an object of class "character"

>So I cheated by doing this:
```{r }

cardiac_diagnosis %>%
     count(care_type_desc, sex_desc, sort=TRUE) %>%
     head(2)

```

>But when I plot it, using the following command:

```{r try2}

 cardiac_diagnosis %>%
     count(care_type_desc, sex_desc, sort=TRUE) %>%
     head(2) %>%
     ggplot()+
     geom_bar(mapping=aes(x=care_type_desc,fill=sex_desc), position="dodge")+
     labs(title="Most common care type, by gender", y="No. of patients")+
     scale_fill_discrete(name='Gender')+
     theme_classic()

```


>I think I am missing something here.

>Is it ok if you can point me to the right direction? 

>Many thanks


# Dennis' response

I am overjoyed to point you in the right direction!  Thanks so much for the question.  Response below


## Attempt 1

This is a great first attempt, trying to use the `filter` to filter out the rows having the most frequent `care_type_desc`.  

```{r  eval=F}

cardiac_diagnosis %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  filter(
         count(care_type_desc, sort=TRUE) %>%
           head(1)
  )

```

And as  you pointed out, it doesn't work. Technically, this is because `filter` expects a boolean condition, something that returns `TRUE` or `FALSE`, usually of the form `x=y`.   If your filter is in that form, it could actually work.  For example, if you knew what your most frequent care type was, for, by example, running this:

```{r}

  cardiac_diagnosis  %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  count(care_type_desc, sort=TRUE) %>%
  head(1) %>%
  select( care_type_desc )

```
We would find that the most frequent care type was `Other care (Acute) inc Qualified Newborn`.  So, we could actually filter by that:

```{r }

cardiac_diagnosis %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  filter( care_type_desc == 'Other care (Acute) inc Qualified Newborn' )

```

This could then be piped into ggplot, to answer your question.  

```{r}

cardiac_diagnosis %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  filter( care_type_desc == 'Other care (Acute) inc Qualified Newborn' ) %>%
  ggplot()+
    geom_bar(mapping=aes(x=care_type_desc,fill=sex_desc), position="dodge")+
    labs(title="Most common care type, by gender", y="No. of patients")+
    scale_fill_discrete(name='Gender')+
    theme_classic()

```
 I can tell that you already know all this!  It is just that you want to do it automatically, without human intervention.  So do I. I show you this because I want to demonstrate 1) `filter` is not wrong, and 2) later, I will show you how to do it with `filter`. It is just that the `filter` approach is a bit complex.

## Try 2 - Cheating by selecting top 2 rows

Your second attempt was as follows:

```{r }
 
cardiac_diagnosis %>%
    count(care_type_desc, sex_desc, sort=TRUE) %>%
    head(2) %>%
    ggplot()+
    geom_bar(mapping=aes(x=care_type_desc,fill=sex_desc), position="dodge")+
    labs(title="Most common care type, by gender", y="No. of patients")+
    scale_fill_discrete(name='Gender')+
    theme_classic()

```

What a good try!  Very innovative.  The problem here, as you demonstrated and can see below, is that your data selection step, the first 3 lines of your code, only return 2 records.  `geom_bar` *counts* the records for you (by default), and it is only counting 2!

```{r }

cardiac_diagnosis %>%
  count(care_type_desc, sex_desc, sort=TRUE) %>%
  head(2) 

```

## Solution 2 - Cheating
Now, if we could get `geom_bar` to use that `n` instead of counting, we would be sweet.  `geom_col` is the not so helpful version of `geom_bar`, it demands a `y` parameter, and it *does not count for you*.  Which is what you want.

```{r }

cardiac_diagnosis %>%
  count(care_type_desc, sex_desc, sort=TRUE) %>%
  head(2)  %>%
  ggplot()+
    geom_col(mapping=aes(x=care_type_desc,y=n, fill=sex_desc), position="dodge")+
    labs(title="Most common care type, by gender", y="No. of patients")+
    scale_fill_discrete(name='Gender')+
    theme_classic()


```

So this makes your second, *cheating*, attempt work.  

## Solution using `inner_join`

But what if you did not want to cheat?  What if you wanted to do that `filter`, and find the actual most frequent care types? 

The trick here is that `inner_join` also filters, *if the thing you are joining has a restricted number of records*.  First, we create a variable holding the most frequent `care_type_desc`.

```{r}

most_frequent_care_type <- 
  cardiac_diagnosis  %>%
    distinct(admission_id, care_type_desc, sex_desc) %>%
    count(care_type_desc, sort=TRUE) %>%
    head(1)

most_frequent_care_type

```

Then, we `inner_join` `cardiac_diagnosis` with this.

```{r}

cardiac_diagnosis %>%
  inner_join( most_frequent_care_type )

```

This has the effect of filtering out cardiac_diagnosis records with only this `care_type_desc`, as demonstrated below.

```{r}

cardiac_diagnosis %>%
  inner_join( most_frequent_care_type ) %>%
  distinct( care_type_desc )

```
We only have this single care_type_desc 

Combining the `inner_join` filter and ggplot, we get the same graph.

```{r}

most_frequent_care_type <- 
  cardiac_diagnosis  %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  count(care_type_desc, sort=TRUE) %>%
  head(1)

cardiac_diagnosis %>%
  inner_join( most_frequent_care_type ) %>%
  ggplot()+
    geom_bar(mapping=aes(x=care_type_desc,fill=sex_desc), position="dodge")+
    labs(title="Most common care type, by gender", y="No. of patients")+
    scale_fill_discrete(name='Gender')+
    theme_classic()

```

## Solution using `filter` and `pluck`

Instead of using `inner_join`, we can use filter directly, if *we know the name of the most frequent care type desc*.  We can extract out the actual name using the `pluck` command.  `pluck` extracts out a single value from a tibble. 


```{r}

most_frequent_care_type_desc <- 
  cardiac_diagnosis  %>%
  distinct(admission_id, care_type_desc, sex_desc) %>%
  count(care_type_desc, sort=TRUE) %>%
  head(1) %>%
  pluck( 1 )

most_frequent_care_type_desc 

cardiac_diagnosis %>%
  filter( care_type_desc == most_frequent_care_type_desc ) %>%
  ggplot()+
  geom_bar(mapping=aes(x=care_type_desc,fill=sex_desc), position="dodge")+
  labs(title="Most common care type, by gender", y="No. of patients")+
  scale_fill_discrete(name='Gender')+
  theme_classic()

```


