---
title:           Money
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          html
---
```{r libraries, message=F, show=F}
library(tidyverse)
library(readxl)
library(RPostgreSQL)

```

# Money Data types

```{r getdata}

con <- dbConnect(dbDriver("PostgreSQL"), 
                 dbname   = 'pbs',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query <- " 
select brand_premium from pbs_item_manufacturer limit 10
 "

money_data <- dbGetQuery( con, query) %>% as_tibble() 

#

```
We get a warning, but the data is retrieved.  The probelm is that the data is of type chararacter (chr), and so we cannot do arithmetic on it. 

We try to convert it to numeric data

```{r convert}

money_data %>% 
  mutate (brand_premium = as.numeric( brand_premium ))

```

That doesn't work very well, because R doesn't know what to do with the $ sign.  

We will use the str_remove function from the stingr package (part of tidyverse) to get rid of the $.



```{r modify_and_convert}

money_data %>% 
  mutate (brand_premium = str_remove( string=brand_premium, pattern='\\$') ) %>%
  mutate (brand_premium = as.numeric( brand_premium ))

```

Note,  because str_remove expects to remove regular expression pattern, and $ has special meaning in the regular expression language, we have to `escape` the $, by putting the \\ in front of it. 

## Try # 2 - do the conversion in the database

Well, that was hard!  Better, because postgres knows what a money type is, we can fix it in SQL - we can get postgres to convert it to a number, using

```{SQL}

select cast(brand_premium as numeric ) from pbs_item_manufacturer

```

Here is the R code to prove it works


```{r take2}

query1 <- " 
select cast( brand_premium as numeric ) as brand_premium from pbs_item_manufacturer limit 10
"

money_data <- dbGetQuery( con, query1) %>% as_tibble() 

money_data


```



