---
title:           Cut Error
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          html
---
```{r message=F, show=F}
library(tidyverse)
library(readxl)

```

# The Question - I've got an error, how do I start to fix it?
1) What is the minimum code needed to reproduce the error?

```{r ,eval=F}

con <- dbConnect(dbDriver("PostgreSQL"), 
                 dbname   = 'vaed_full',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query_diag <- " 
select diag_code, admission_id, age_years, los, sameday_id, drg, sex_desc, diag_short_desc, sep_mode_desc, care_type_desc
from admission JOIN sex using (sex_id)
join admission_diagnosis using (admission_id)
join diagnosis_desc using (diag_code)
JOIN separation_mode USING (sep_mode_id)
JOIN care_type USING (care_type_id)
WHERE diag_code='a370';
 "

pertussis_diagnosis <- dbGetQuery( con, query_diag ) %>% as_tibble() 
#

query_proc <- " 
select procedure_code, admission_id, age_years, los, sameday_id, drg, sex_desc, procedure_short_desc, block_desc, sep_mode_desc, care_type_desc
from admission JOIN sex using (sex_id)
join admission_procedure using (admission_id)
join procedure_desc using (procedure_code)
join block using (block)
JOIN separation_mode USING (sep_mode_id)
JOIN care_type USING (care_type_id)
WHERE block='1916' "

#
Pertussis_proc <- dbGetQuery( con, query_proc ) %>% as_tibble() 
#

library('readxl')
library('janitor')

population <- 
  read_excel('populationVictoria_2014_long (1).xlsx') %>%
  clean_names()

pertussis_summary <- 
  pertussis_diagnosis %>%
  mutate( ten_year_age_group = cut( age_years, breaks=10 * (0:12 ))) %>%
  count( ten_year_age_group, sex_desc ) 

pertussis_summary
```

Error in dbConnect(dbDriver("PostgreSQL"), dbname = "vaed_full", host = "himsql7.latrobe.edu.au",  :
could not find function "dbConnect"

This code does not run because it does not load the libraries.  More importantly, it should run standalone, all by itself.

First, delete lines unrelated to the error.   The error came when you were trying to count pertussis_diagnosis. The original code had sections dealing with procedures, and population.  When trying to find out what is causing an error, simplify it as much as possible.

The minimum remaining code that produces the error is below.

We go through this code line by line, and verify that each line is doing what we expect.

We can either match it against other code that we know that is correct, or make sure that it does what it is supposed to do, or even better yet, both!

For example, the first 2 lines, the library lines, are they identical to the other code that works?

The part where we make query_diag, does that work?  Does the query_diag get the correct query inside?  Does the SQL query work in pgadmin?

Then, we go get pertussis_diagnosis tibble.  Does that line work?  Does pertussis_diagnosis get filled, after that line is traversed?

Finally, we execute the pipeline, where we take pertussis_diagnosis, and mutate it.  The error comes when we try to mutate it, it says that it cannot find age_years.  Does it contain age_years?  Go check...


```{r ,eval=F}

library(tidyverse)
library(RPostgreSQL)

con <- dbConnect(dbDriver("PostgreSQL"), 
                 dbname   = 'vaed_full',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query_diag <- " 
select diag_code, admission_id, age_years, los, sameday_id, drg, sex_desc, diag_short_desc, sep_mode_desc, care_type_desc
from admission JOIN sex using (sex_id)
join admission_diagnosis using (admission_id)
join diagnosis_desc using (diag_code)
JOIN separation_mode USING (sep_mode_id)
JOIN care_type USING (care_type_id)
WHERE diag_code='a370';
"

pertussis_diagnosis <- dbGetQuery( con, query_diag ) %>% as_tibble() 
#

pertussis_summary <- 
  pertussis_diagnosis %>%
  mutate( ten_year_age_group = cut( age_years, breaks=10 * (0:12 ))) %>%
  count( ten_year_age_group, sex_desc ) 

```

