---
title:           Money
author:          Dennis Wollersheim
date:            27.05.2019
linkcolor:       cyan
output:          html
---
```{r libraries, message=F, show=F}
library(tidyverse)
library(readxl)
library(RPostgreSQL)

```

# Money Data types

```{r getdata}

con <- dbConnect(dbDriver("PostgreSQL"),
                 dbname   = 'pbs',
                 host = "himsql7.latrobe.edu.au", port = 5432,
                 user = "dewollershei-test", password = "healthGuru")

query <- "
select benpaid from mbs limit 10
 "

money_data <- dbGetQuery( con, query) %>% as_tibble()

#

```
We get a warning, but the data is retrieved.  The probelm is that the data is of type chararacter (chr), and so we cannot do arithmetic on it.

We try to convert it to numeric data

```{r convert}

money_data %>%
  mutate (benpaid = as.numeric( benpaid ))

```

That doesn't work very well, because R doesn't know what to do with the $ signs and commas.

We will use the str_remove_all function from the stingr package (part of tidyverse) to get rid of them.



```{r modify_and_convert}

money_data %>%
  mutate (benpaid = str_remove_all( string=benpaid, pattern='[$,]') ) %>%
  mutate (benpaid = as.numeric( benpaid ))

```

## Try # 2 - do the conversion in the database

Well, that was hard!  Better, because postgres knows what a money type is, we can fix it in SQL - we can get postgres to convert it to a number, using

```{sql connection=con}

select cast(benpaid as numeric ) from mbs limit 10

```

Here is the R code to prove it works


```{r take2}

query1 <- "
select cast( benpaid as numeric ) as benpaid from mbs limit 10
"

money_data <- dbGetQuery( con, query1) %>% as_tibble()

money_data


```



