
---
title:           
author:          Dennis Wollersheim 
date:            27.05.2019
linkcolor:       cyan
output:          html
---
```{r,warning=FALSE,message=FALSE,echo=FALSE}
# load libraries
library(knitr)
opts_chunk$set(warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)

```


# The Question - Top N filtering

> I want to compare the age distribution between procedures, but there are too many procedures.  How can I get the age distribution for the top 4 procedures only?

To demonstrate, I will use the `cardiac_procedures` dataset.


## Setup Load Libraries and Data

```{r}

library(tidyverse)
cardiac_procedures <- read_csv("cardiac_procedures.csv") 

```

## The problem
THis graph demonstrates the problem.  There are too many categories.

```{r}

cardiac_procedures %>%
  ggplot( aes( age_years ,fill = procedure_short_desc )) +
  geom_histogram(bins=10)


```


## Finding the top 4 procedures
It is easy to find the top 4 procedures, by first counting how many admissions for each procedure, sorting them in descending order, and throwing away everything from #5 onwards.

```{r }

cardiac_procedures %>%
  count(procedure_short_desc, sort = TRUE) %>%
  head( 4)

```

The problem is that now that we have the top 4 procedures we no longer have the other admission details, e.g. age.  

# The full solution

We do need these 4 procedures, but we need to use them in conjunction with the full dataset, to use them to filter the `cardiac_procedures` dataset, selecting the row containing these 4 procedures.  Finally, we can graph that filtered dataset.

First, store away the top 4 procedures in a tibble called `top_4_cardiac_procedures`.

```{r }

top_4_cardiac_procedures <-
  cardiac_procedures %>%
    count(procedure_short_desc, sort = TRUE) %>%
    head( 4)

```

Then, again, use `inner_join` to do the filtering, joining `top_4_cardiac_procedures` to the full `cardiac_procedures` dataset, and pipe the resulting filtered dataset into ggplot.

```{r}

cardiac_procedures %>%
  inner_join( top_4_cardiac_procedures, by='procedure_short_desc') %>%
  ggplot( aes( age_years ,fill = procedure_short_desc )) +
  geom_histogram(bins=10)

```
