---
title: "Tutorial 7 - Top N query"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
---


# Top N queries in R


## Load Libraries


```{r,warning=FALSE,message=FALSE,echo=TRUE}

# load libraries
library(knitr)
opts_chunk$set(warning=F,message=F,fig.width = 11,fig.height = 5,cache=F, echo=T)
library(tidyverse)
library("RPostgreSQL")
library("forcats")

```
## Setup Connection

```{r database_setup, echo=TRUE, warnings=FALSE, messages=FALSE}

# create connection to the database
con <- dbConnect(dbDriver("PostgreSQL"),
				 dbname   = 'vaed_full',
				 host = "himsql7.latrobe.edu.au", port = 5432,
				 user = "dewollershei-test", password = "healthGuru")
#
# this is the query to get all tne pneumonia diagnois.  the 3 commented lines would also finn the principal procedure
```

## Setup Query
```{r setup_query}
query = "
SELECT ad_morbidity.position, ad_morbidity.diag_code, diag_short_desc, admission_id, age_years, los
FROM admission JOIN sex USING (sex_id)
JOIN admission_diagnosis ad_morbidity USING (admission_id)
JOIN diagnosis_desc USING (diag_code)
JOIN admission_diagnosis ad_pneumonia USING (admission_id)
WHERE ad_pneumonia.diag_code = 'J189'
AND ad_morbidity.position = 1
"

```

## Get data from database

we want to get rid of records with the J189, because we want the other morbidities

```{r}

dbGetQuery( con, query ) %>%
  as_tibble() %>%
  filter( diag_code != 'J189') %>%
  { . } -> morbidities

```

# What are the morbidites that have an additional diagnosis of pneumonia

note, there are 1400 morbidities

```{r what morbitidies }

morbidities %>%
  count( diag_short_desc, sort=TRUE)

morbidities %>%
  ggplot( aes( x=diag_short_desc )) +
  geom_bar()

```



# What can we do about this?

We can lump together the smaller categories into an 'other' category

```{r top_5}

morbidities %>%
  filter( diag_code != 'J189') %>%
  mutate( diag_short_desc = fct_lump_prop(diag_short_desc,0.05)) %>%
  count( diag_short_desc, sort=TRUE)

morbidities %>%
  filter( diag_code != 'J189') %>%
  mutate( diag_short_desc = fct_lump_n(diag_short_desc,10)) %>%
  count( diag_short_desc, sort=TRUE)
```

#  Graphing it

```{r}

morbidities %>%
  filter( diag_code != 'J189') %>%
  mutate( diag_short_desc = fct_lump_n(diag_short_desc,10)) %>%
  filter(diag_short_desc != 'Other') %>%
  ggplot( aes( x=diag_short_desc, fill=diag_short_desc )) +
  geom_bar()+
  coord_flip()

```

# Graph is better if ordered by largest values

We need to count the morbidities first, so we know how to order them.  Once they are counted, we use geom_col to graph them

```{r}

morbidities %>%
  filter( diag_code != 'J189') %>%
  mutate( diag_short_desc = fct_lump_n(diag_short_desc,10)) %>%
  filter(diag_short_desc != 'Other') %>%
  count( diag_short_desc) %>%
  mutate( diag_short_desc = fct_reorder(diag_short_desc, n)) %>%
  ggplot( aes( x=diag_short_desc, y = n, fill=diag_short_desc )) +
  geom_col()+
  coord_flip()



```



